---
title: "Exploring Potential Environmental Racism in Cancer Alley, Louisiana"
author: "Ashley Weaver"
date: "May 2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    code_folding: hide
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Introduction

Cancer Alley, Louisiana is an area along the Mississippi River between Baton Rouge and New Orleans. It derived its notorious nickname, Cancer Alley, from the abundance of oil refineries and petrochemical companies along the river. It is already known that this area has an unusually high rate of cancers, which is attributed to the air and water contamination by industry. Environmental Racism in Cancer Alley, Louisiana has previously been explored but mostly in terms of cancer. This study will be exploring the other potential health risks that people living in Cancer Alley experience compared to the rest of Louisiana. This study will also look at other factors that could explain differences in quality of health by county.

The state of Louisiana is known for having very limited environmental regulations and therefore very high levels of pollution. Louisiana was ranked by US News as the worst pollution state in terms of exposure to pollution and related health risks. Additionally, there are numerous articles documenting the lack of enforcement of federal clean air, clean water, and hazardous waste laws. The lack of environmental regulation is part of the reason for the abundance of industrial sites in the state. High levels of industry and low levels of environmental protection have led to adverse health impacts throughout the state. Previous studies have documented the effects of these industrial sites on the number of cancer cases in the state.

There are measures of increased cancer risk and racial data for some of the counties along what is known as Cancer Alley. For instance, the lifetime risk of cancer in St. John Baptist is 800 times higher than an average american (Pasley 2020). Nationally, there has been a 16% decrease in toxic releases by plants over the last 30 year, but in Cancer Alley reported toxic releases grew by 25%. It is also noted that these oil refinery and chemical plants are located next to African American communities along the Mississippi River (Pasley 2020). 

This study will be focused on exploring race and health data in the state of Louisiana to assess whether there is a link between health problems and race, particularly in Cancer Alley. It is important to note that this study is limited to looking at data by counties, though Cancer Alley itself is not entire counties. Populations that are closer to the industries and the Mississippi River (where these industries dump waste) are more likely to be impacted in terms of health. Since the study looks at health in Louisiana at a county level, it is possible that the adverse health impacts to communities on the river will not be reflected by the county-wide data.



## Methods

Three difference R packages were utilized to create the following analysis. The dplyr package was used extensively to create new descriptive variables and reorganize and restructure data frames to facilitate the analysis. The ggplot2 package was used in order to visualize the analysis through the creation of a variety of different plot types. Finally, the tidyr package was used to combine data from the race and health datasets so that they could be analyzed on the same plot.

The first portion of my analysis will focus on race data for counties that are considered to be part of Cancer Alley. These counties include: East Baton Rouge, West Baton Rouge, Ascension, St. John the Baptist, Iberville, St. Charles, Jefferson, and Orleans. My analysis will explore the differences in diversity of these areas compared to one another and the state as a whole. This will be followed by a discussion of the limitation of my racial analysis given the structure of the data. 

The second part of my analysis will look at race and health data in the entire state of Louisiana. This portion of the analysis will focus on finding correlation between poor health and percentage of minorities in a county. This will include a scatter plot that examines health as a function of percentage of minorities in the community. This will give a more holistic view of Louisiana's health by race. This will be followed by theoretical qq-plots to assess the normality of the data and isolate particular counties in Louisiana that appear to be in worse health. The health data will then be compared to other states in the United States to assess Louisiana's statewide health compared to national health.

The third part of my analysis will focus primarily on the health in the counties that are part of Cancer Alley. This portion of the analysis uses a box plot to compare the percentage poor health in the counties of Cancer Valley, Louisiana to percentage poor health in the state and country as a whole.  

The final portion of my analysis will review other factors that could be contributing to adverse health effects. This will involve an evaluation of the potential impact of lack of health insurance on Louisiana. This will be demonstrated through box plots that will compare the lack of insurance by age in the US and lack of insurance by age in Louisiana. Comparing the difference between National lack of insurance to just state lack of insurance in Louisiana will give us a better idea of if insurance (or lack thereof) can be used to explain the health disparity between most U.S. States and Louisiana. Next, this section will explore potential causes of obesity in Louisiana, which is shown in section 2 ("Comparing Race and Health Data for the Entire State of Louisiana") to have the second highest rate of obesity in the US. This analysis will involve the use of scatterplots to compare obesity in Louisiana to access to healthy foods and physical inactivity.



## Results and Discussion

### Cancer Alley Race Data

```{r packages, include = FALSE}

library(dplyr)
library(ggplot2)
library(tidyr)

```

